A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-12-05 tex/codatadp.tex
more

For codatatype, the dual notion is that of a coinduction principle that
depends on witnesses $R$ that are required to be bisimulations:
%
\[
\begin{aligned}[t]
\text{Coinduction:}\quad
\AXC{\strut$\begin{gathered}\textstyle \bigwedge_i R_i[v_i, w_i] \\[-\jot]\textstyle
\bigwedge_i \forall v\; w.\;\, R_i[v, w] \longrightarrow
  \bigwedge_j \const{d}_j(v) \teq \const{d}_j(w)
  \mathrel\land
  \const{d}_j(v) \longrightarrow \bigwedge_k \const{s}_{i\negvthinspace j}^{\,k}(v) \sim \const{s}_{i\negvthinspace j}^{\,k}(w)
\end{gathered}$}
\UIC{\strut$\bigwedge_i v_i \teq w_i$}
\DP
\end{aligned}
\]
where

This axiom schema may look intimidating, but what it

For codatatypes

    * datatypes:
      * induction axiom (schema)
        * consequence: no infinite objects (in particular, no cyclic objects)

      * when looking at the universal ground (?) theory:
        enough to consider acyclicity -- no way to specify infinite objects
        otherwise

    * codatatypes:
      * coinduction axiom (schema)
        * consequence: infinite objects are allowed, but
          two objects yielding the same observations must be equal

@@

  * more evaluation
    * use it in a new Nitpick-like tool
  * recursive and corecursive functions
    (and rewriting)
  * extensions, perhaps straightforward, to polymorphism and nested
    (co)recursion

>>>>

For codatatype, the dual notion is that of a coinduction principle that
depends on witnesses $R$ that are required to be bisimulations:
%
\[
\begin{aligned}[t]
\text{Coinduction:}\quad
\AXC{\strut$\begin{gathered}\textstyle \bigwedge_i R_i[v_i, w_i] \\[-\jot]\textstyle
\bigwedge_i \forall v\; w.\;\, R_i[v, w] \longrightarrow
  \bigwedge_j \const{d}_j(v) \teq \const{d}_j(w)
  \mathrel\land
  \const{d}_j(v) \longrightarrow \bigwedge_k \const{s}_{i\negvthinspace j}^{\,k}(v) \sim \const{s}_{i\negvthinspace j}^{\,k}(w)
\end{gathered}$}
\UIC{\strut$\bigwedge_i v_i \teq w_i$}
\DP
\end{aligned}
\]
where $x \sim y$ denotes either $R_i[x, y]$ if there exists some $i$ such that
this is type correct or $x \teq y$ otherwise.

FIXME: Find out what ensures that codatatypes have infinite values.

This axiom schema captures the idea that two values that yield the same
observations must be equal, where the observations are made by using the
selectors and discriminators.

% TODO: where does this go?
%      * when looking at the universal ground (?) theory:
%        enough to consider acyclicity -- no way to specify infinite objects
%        otherwise

%    * codatatypes:
%      * coinduction axiom (schema)
%        * consequence: infinite objects are allowed, but
%          two objects yielding the same observations must be equal

@@

  * more evaluation
    * use it in a new Nitpick-like tool
  * recursive and corecursive functions
    (and rewriting)
  * extensions, perhaps straightforward, to polymorphism and nested
    (co)recursion
  * quantifiers?
